\bf Model Visualization Techniques for a Social Network Model

### 2017-03-12 paper.Rnw
write write write write write

@@

The guiding principles XXX cite XXX of model visualization are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foreground, stochastic actor-oriented models will be better understood and more accessible to social network analysts.

>>>>

Models for dynamic social networks are extremely important because of how realistic they are. Social networks do not form spontaneously; they evolve over time. Ties are added and removed, and new actors can join the network. Modeling the underlying mechanisms that create network changes over time is very complex but also has the potential to uncover hidden truths.

\par One set of models for dynamic social networks are the stochastic actor-oriented models (SAOMs) of \citet{saompaper}. These models are uniquely different from other social network models because they incorporate network statistics and actor statistics into the model, whereas most other models just use the network statistics to determine change. Allowing the actor-level statistics to change the structure of the network leads to a more intuitive model for social network change. In the "real-world" we would expect that people with common interests would be more likely to form relationships, and that intuition is no longer ignored in modeling with SAOMs.

@@

The guiding principles of model visualization from \citet{modelvis} are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foreground, stochastic actor-oriented models will be better understood and more accessible to social network analysts.
